Lattices of acyclic pipe dreams

### 2023-03-16 acyclicFacets/draft/latticePipeDreams.tex
title subsection

\begin{example}
We have represented five examples in \cref{fig:subwordComplex1,fig:subwordComplex2,fig:subwordComplex3,fig:subwordComplex4,fig:subwordComplex5}.
The last three use~$\omega = \wo$ and are spherical subword complexes (their brick polytopes are represented in \cref{fig:brickPolytope1,fig:brickPolytope2,fig:brickPolytope3}), while the first two do not.
We use the representation of the facets as pseudoline arrangements on sorting networks, see \cref{exm:sortingNetworks}.
%
\begin{figure}[p]
	\centerline{\includegraphics[scale=.7]{subwordComplex4}}
	\caption{The subword complex equivalence (left) and the acyclic increasing flip graph (right) for~$Q = \tau_2 \tau_3 \tau_1 \tau_3 \tau_2 \tau_1 \tau_2 \tau_3 \tau_1$ and~$w = \tau_2 \tau_3 \tau_1 \tau_2 \tau_3 = 3421$ in type~$A_3$. Note that the sets~$\linearExtensions(I)$ (blue bulles) are not always intervals.}
	\label{fig:subwordComplex4}
\end{figure}
%
\begin{figure}[p]
	\centerline{\includegraphics[scale=.7]{subwordComplex5}}
	\caption{The subword complex equivalence (left) and the acyclic increasing flip graph (right) for~$Q = \tau_3 \tau_4 \tau_1 \tau_3 \tau_2 \tau_4 \tau_1 \tau_2$ and~$w = \tau_3 \tau_4 \tau_1 \tau_2 = 24153$ in type~$A_4$. Note that the sets~$\linearExtensions(I)$ (blue bubbles) are not always subsets of, and may even not meet~$[e,\omega]$ (green bubble).}
	\label{fig:subwordComplex5}
\end{figure}
%
\hvFloat[floatPos=p, capWidth=h, capPos=r, capAngle=90, objectAngle=90, capVPos=c, objectPos=c]{figure}
{\includegraphics[scale=.6]{subwordComplex1}}
{For~$Q = \tau_2 \tau_1 \tau_3 \tau_2 \tau_1 \tau_3 \tau_2 \tau_1 \tau_3$ and~$\omega = \wo = 4321$ in type~$A_3$, the subword complex equivalence is a lattice congruence (left) so that the increasing flip poset is a lattice (right). See also~\cref{fig:increasingFlipPosets}\,(left) It is actually the $\tau_2\tau_1\tau_3$-Cambrian lattice in type~$A_3$. The blue bubbles represent the classes of the pipe dream congruence. For each facet of the subword complex, we have represented the corresponding pseudoline arrangement on the sorting network, and the corresponding contact graph. Note that all facets are acyclic in this case.}
{fig:subwordComplex1}
%
\hvFloat[floatPos=p, capWidth=h, capPos=r, capAngle=90, objectAngle=90, capVPos=c, objectPos=c]{figure}
{\includegraphics[scale=.7]{subwordComplex2}}
{For~$Q = \tau_2 \tau_3 \tau_1 \tau_3 \tau_2 \tau_1 \tau_2 \tau_3 \tau_1$ and~$\omega = \wo = 4321$ in type~$A_3$, the subword complex equivalence is not a lattice congruence (left) but the increasing flip poset is a lattice (right). The blue bubbles represent the classes of the pipe dream congruence. For each facet of the subword complex, we have represented the corresponding pseudoline arrangement on the sorting network, and the corresponding contact graph. Note that all facets are acyclic in this case.}
{fig:subwordComplex2}
%
\hvFloat[floatPos=p, capWidth=h, capPos=r, capAngle=90, objectAngle=90, capVPos=c, objectPos=c]{figure}
{\includegraphics[scale=.7]{subwordComplex3}}
{For~$Q = \tau_1 \tau_2 \tau_3 \tau_2 \tau_1 \tau_2 \tau_3 \tau_2 \tau_1$ and~$\omega = \wo = 4321$ in type~$A_3$, the subword complex equivalence is not a lattice congruence (left) and the increasing flip poset is not a lattice (right). See also~\cref{fig:increasingFlipPosets}\,(right). The blue bubbles represent the classes of the pipe dream congruence. For each facet of the subword complex, we have represented the corresponding pseudoline arrangement on the sorting network, and the corresponding contact graph. Note that all facets are acyclic in this case.}
{fig:subwordComplex3}
\end{example}

@@

\begin{example}
We have illustrated in \cref{fig:brickPolytope1,fig:brickPolytope2,fig:brickPolytope3} the brick polytopes of the subword complexes represented in \cref{fig:subwordComplex1,fig:subwordComplex2,fig:subwordComplex3}.
Note that the oriented graph (from bottom to top) defines a lattice in \cref{fig:brickPolytope1,fig:brickPolytope2} but not in \cref{fig:brickPolytope3}.
%
\begin{figure}[p]
	\centerline{\input{figures/brickPolytope1}}
	\caption{The brick polytope of~$\subwordComplex(Q,\wo)$ where~$Q = \tau_2 \tau_1 \tau_3 \tau_2 \tau_1 \tau_3 \tau_2 \tau_1 \tau_3$ in type~$A_3$. Its graph, oriented from bottom to top is the lattice of \cref{fig:increasingFlipPosets}\,(left) and \cref{fig:subwordComplex1}. This is actually the $c$-associahedron and the $c$-Cambrian lattice for the Coxeter element~$c = \tau_2\tau_1\tau_3$ of type~$A_3$.}
	\label{fig:brickPolytope1}
\end{figure}
%
\begin{figure}[p]
	\centerline{\input{figures/brickPolytope2}}
	\caption{The brick polytope of~$\subwordComplex(Q,\wo)$ where~$Q = \tau_2 \tau_3 \tau_1 \tau_3 \tau_2 \tau_1 \tau_2 \tau_3 \tau_1$ in type~$A_3$. Its graph, oriented from bottom to top is the lattice of \cref{fig:subwordComplex2}, but it is not obtained as a lattice quotient of the weak order.}
	\label{fig:brickPolytope2}
\end{figure}
%
\begin{figure}[p]
	\centerline{\input{figures/brickPolytope3}}
	\caption{The brick polytope of~$\subwordComplex(Q,\wo)$ where~$Q = \tau_1 \tau_2 \tau_3 \tau_2 \tau_1 \tau_2 \tau_3 \tau_2 \tau_1$ in type~$A_3$. Its graph, oriented from bottom to top is the poset of \cref{fig:increasingFlipPosets}\,(right) and \cref{fig:subwordComplex3}, which is not a lattice.}
	\label{fig:brickPolytope3}
\end{figure}
\end{example}

@@

\subsection{The sweeping algorithm}
\label{subsec:sweepingAlgorithmSubwordComplexes}

>>>>

\begin{example}
We have represented five examples in \cref{fig:subwordComplex1,fig:subwordComplex2,fig:subwordComplex3,fig:subwordComplex4,fig:subwordComplex5}.
The last three use~$\omega = \wo$ and are spherical subword complexes (their brick polytopes are represented in \cref{fig:brickPolytope1,fig:brickPolytope2,fig:brickPolytope3}), while the first two do not.
We use the representation of the facets as pseudoline arrangements on sorting networks, see \cref{exm:sortingNetworks}.
%
\begin{figure}[p]
	\centerline{\includegraphics[scale=.7]{subwordComplex4}}
	\caption{The subword complex equivalence (left) and the acyclic increasing flip graph (right) for~$Q = \tau_2 \tau_3 \tau_1 \tau_3 \tau_2 \tau_1 \tau_2 \tau_3 \tau_1$ and~$w = \tau_2 \tau_3 \tau_1 \tau_2 \tau_3 = 3421$ in type~$A_3$. Note that the sets~$\linearExtensions(I)$ (blue bulles) are not always intervals.}
	\label{fig:subwordComplex4}
\end{figure}
%
\begin{figure}[p]
	\centerline{\includegraphics[scale=.7]{subwordComplex5}}
	\caption{The subword complex equivalence (left) and the acyclic increasing flip graph (right) for~$Q = \tau_3 \tau_4 \tau_1 \tau_3 \tau_2 \tau_4 \tau_1 \tau_2$ and~$w = \tau_3 \tau_4 \tau_1 \tau_2 = 24153$ in type~$A_4$. Note that the sets~$\linearExtensions(I)$ (blue bubbles) are not always subsets of, and may even not meet~$[e,\omega]$ (green bubble).}
	\label{fig:subwordComplex5}
\end{figure}
%
\hvFloat[floatPos=p, capWidth=h, capPos=r, capAngle=90, objectAngle=90, capVPos=c, objectPos=c]{figure}
{\includegraphics[scale=.55]{subwordComplex1}}
{For~$Q = \tau_2 \tau_1 \tau_3 \tau_2 \tau_1 \tau_3 \tau_2 \tau_1 \tau_3$ and~$\omega = \wo = 4321$ in type~$A_3$, the subword complex equivalence is a lattice congruence (left) so that the increasing flip poset is a lattice (right). See also~\cref{fig:increasingFlipPosets}\,(left) It is actually the $\tau_2\tau_1\tau_3$-Cambrian lattice in type~$A_3$. The blue bubbles represent the classes of the pipe dream congruence. For each facet of the subword complex, we have represented the corresponding pseudoline arrangement on the sorting network, and the corresponding contact graph. Note that all facets are acyclic in this case.}
{fig:subwordComplex1}
%
\hvFloat[floatPos=p, capWidth=h, capPos=r, capAngle=90, objectAngle=90, capVPos=c, objectPos=c]{figure}
{\includegraphics[scale=.65]{subwordComplex2}}
{For~$Q = \tau_2 \tau_3 \tau_1 \tau_3 \tau_2 \tau_1 \tau_2 \tau_3 \tau_1$ and~$\omega = \wo = 4321$ in type~$A_3$, the subword complex equivalence is not a lattice congruence (left) but the increasing flip poset is a lattice (right). The blue bubbles represent the classes of the pipe dream congruence. For each facet of the subword complex, we have represented the corresponding pseudoline arrangement on the sorting network, and the corresponding contact graph. Note that all facets are acyclic in this case.}
{fig:subwordComplex2}
%
\hvFloat[floatPos=p, capWidth=h, capPos=r, capAngle=90, objectAngle=90, capVPos=c, objectPos=c]{figure}
{\includegraphics[scale=.65]{subwordComplex3}}
{For~$Q = \tau_1 \tau_2 \tau_3 \tau_2 \tau_1 \tau_2 \tau_3 \tau_2 \tau_1$ and~$\omega = \wo = 4321$ in type~$A_3$, the subword complex equivalence is not a lattice congruence (left) and the increasing flip poset is not a lattice (right). See also~\cref{fig:increasingFlipPosets}\,(right). The blue bubbles represent the classes of the pipe dream congruence. For each facet of the subword complex, we have represented the corresponding pseudoline arrangement on the sorting network, and the corresponding contact graph. Note that all facets are acyclic in this case.}
{fig:subwordComplex3}
\end{example}

@@

\begin{example}
We have illustrated in \cref{fig:brickPolytope1,fig:brickPolytope2,fig:brickPolytope3} the brick polytopes of the subword complexes represented in \cref{fig:subwordComplex1,fig:subwordComplex2,fig:subwordComplex3}.
Note that the oriented graph (from bottom to top) defines a lattice in \cref{fig:brickPolytope1,fig:brickPolytope2} but not in \cref{fig:brickPolytope3}.
%
\begin{figure}[p]
	\centerline{\input{figures/brickPolytope1}}
	\caption{The brick polytope of~$\subwordComplex(Q,\wo)$ where~$Q = \tau_2 \tau_1 \tau_3 \tau_2 \tau_1 \tau_3 \tau_2 \tau_1 \tau_3$ in type~$A_3$. Its graph, oriented from bottom to top is the Hasse diagram of the lattice of \cref{fig:increasingFlipPosets}\,(left) and \cref{fig:subwordComplex1}. This is actually the $c$-associahedron and the $c$-Cambrian lattice for the Coxeter element~$c = \tau_2\tau_1\tau_3$ of type~$A_3$.}
	\label{fig:brickPolytope1}
\end{figure}
%
\begin{figure}[p]
	\centerline{\input{figures/brickPolytope2}}
	\caption{The brick polytope of~$\subwordComplex(Q,\wo)$ where~$Q = \tau_2 \tau_3 \tau_1 \tau_3 \tau_2 \tau_1 \tau_2 \tau_3 \tau_1$ in type~$A_3$. Its graph, oriented from bottom to top is the Hasse diagram of the lattice of \cref{fig:subwordComplex2}, but it is not obtained as a lattice quotient of the weak order.}
	\label{fig:brickPolytope2}
\end{figure}
%
\begin{figure}[p]
	\centerline{\input{figures/brickPolytope3}}
	\caption{The brick polytope of~$\subwordComplex(Q,\wo)$ where~$Q = \tau_1 \tau_2 \tau_3 \tau_2 \tau_1 \tau_2 \tau_3 \tau_2 \tau_1$ in type~$A_3$. Its graph, oriented from bottom to top is the Hasse diagram of the poset of \cref{fig:increasingFlipPosets}\,(right) and \cref{fig:subwordComplex3}, which is not a lattice.}
	\label{fig:brickPolytope3}
\end{figure}
\end{example}

@@

\subsection{Sweeping algorithm}
\label{subsec:sweepingAlgorithmSubwordComplexes}
